\documentclass[10pt]{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multicol}
\setlength{\parindent}{0pt}
\pagestyle{empty}

\begin{document}

% --------------------------------------------------------------------
% Network Architecture and Layers
% --------------------------------------------------------------------
\section*{Network Architecture and Layers}
\begin{multicols}{2}
\textbf{Layer Model:}
\begin{itemize}
    \item Application: End-user services (e.g., HTTP, FTP)
    \item Transport: Process-to-process data transfer (e.g., TCP, UDP)
    \item Network: Routing of datagrams (e.g., IP)
    \item Link: Data transfer between adjacent nodes (e.g., Ethernet, WiFi)
    \item Physical: Transmission of raw bits (e.g., cabling, RF)
\end{itemize}

\textbf{Encapsulation:}
\begin{verbatim}
Application Data (M)
    Transport Header (H_t)
    Network Header (H_n)
    Link Header (H_l)
\end{verbatim}

\columnbreak

\textbf{Layer Responsibilities:}
\begin{itemize}
    \item \textbf{Application:} Network process to application. (e.g., HTTP, DNS)
    \item \textbf{Transport:} Reliable/unreliable data transfer. (e.g., TCP, UDP).
    \item \textbf{Network:} Logical addressing, routing. (e.g., IP).
    \item \textbf{Link:} Physical addressing, frame transmission. (e.g., Ethernet).
    \item \textbf{Physical:} Media, signal, and binary transmission.
\end{itemize}
\end{multicols}

% --------------------------------------------------------------------
% Key Protocols (HTTP, DNS, TCP/UDP, IP)
% --------------------------------------------------------------------
\section*{Key Protocols}
\textbf{HTTP:} HyperText Transfer Protocol; used for transmitting web pages.

\textbf{DNS:} Domain Name System; translates domain names to IP addresses.

\textbf{TCP/UDP:}
\begin{itemize}
    \item \textbf{TCP:} Connection-oriented, reliable, flow control, congestion control.
    \item \textbf{UDP:} Connectionless, unreliable, low overhead.
\end{itemize}

\textbf{IP:} Internet Protocol; handles routing of packets between source and destination.

\begin{multicols}{2}
\textbf{TCP Services:}
\begin{itemize}
    \item Connection-Oriented: Handshake; prepares for data exchange.
    \item Reliable Data Transfer: Ensures error-free, in-order delivery.
\end{itemize}

\textbf{UDP Services:}
\begin{itemize}
    \item Connectionless: No handshake before data transfer.
    \item Unreliable: No guarantees on message delivery.
\end{itemize}
\end{multicols}

% --------------------------------------------------------------------
% Packet vs. Circuit Switching
% --------------------------------------------------------------------
\section*{Packet vs. Circuit Switching}
\textbf{Packet Switching:}
\begin{itemize}
    \item Data sent in packets.
    \item Each packet routed independently.
    \item Pros: Efficient, shared links.
    \item Cons: Variable delays.
\end{itemize}

\textbf{Circuit Switching:}
\begin{itemize}
    \item Dedicated path for duration of the connection.
    \item Pros: Consistent E2E delay.
    \item Cons: Inefficient for bursty traffic.
\end{itemize}

\textbf{Diagrams:}
\begin{verbatim}
Packet:
    [Host] -- [Router] -- [Router] -- [Host]
Circuit:
    [Host] == [Router] == [Router] == [Host] (dedicated)
\end{verbatim}

% --------------------------------------------------------------------
% Performance Metrics (Delay, Throughput, Packet Loss)
% --------------------------------------------------------------------
\section*{Performance Metrics}
Types of Delays:
\begin{itemize}
    \item Processing Delay: Time for router to process the packet header.
    \item Queuing Delay: Time waiting at the queue for transmission.
    \item Transmission Delay: \( L/R \) (packet length / transmission rate).
    \item Propagation Delay: \( d/s \) (distance / propagation speed).
\end{itemize}

\textbf{Throughput:} Rate at which data is successfully transferred (\( \text{bits/sec} \)).

\textbf{Packet Loss:} Packets that don't reach the destination.

\textbf{Formula:}
\[
\text{Total Delay} = d_{\text{proc}} + d_{\text{queue}} + d_{\text{trans}} + d_{\text{prop}}
\]

% --------------------------------------------------------------------
% Web Technologies and CDNs
% --------------------------------------------------------------------
\section*{Web Technologies and CDNs}
\textbf{Web Technologies:}
\begin{itemize}
    \item HTTP(S), HTML, CSS, JavaScript.
    \item Client-server model.
\end{itemize}

\textbf{CDN:} Content Delivery Network; distributes content to servers closer to end-users to reduce latency.

\textbf{Diagram:}
\begin{verbatim}
[Client] -- [Edge Server]
  \/ CDN \/       ||
[Data Center] -- [Origin Server]
\end{verbatim}

% --------------------------------------------------------------------
% Video Streaming Concepts
% --------------------------------------------------------------------
\section*{Video Streaming Concepts}
\textbf{Streaming Types:}
\begin{itemize}
    \item \textbf{Live:} Real-time broadcasting.
    \item \textbf{On-Demand:} Pre-recorded content.
\end{itemize}

\textbf{Important Metrics:}
\begin{itemize}
    \item Bandwidth
    \item Buffering
    \item Bitrate (e.g., 360p, 720p, 1080p)
\end{itemize}

\textbf{Example: Bitrate Calculation}
\[
\text{Total size} = \text{Bitrate} \ \times \ \text{Duration}
\]
% (e.g., \( 2 \, \text{Mbps} \times 60 \, \text{sec} = 15 \, \text{MB})
\end{document}